---
title: "Analysis for CFTR-Pneumonia"
author: "D. Erik Boonstra and Joe E. Cavanaugh"
output: html_document
---

# Purpose
In this project, we are trying to determine if carriers have a higher global trend and larger seasonal trend peaks than controls.

# Data
The data ranges from 2008 to 2019 because:

- excluding 2020 and above due to the possibility of classifying the flu as COVID, and
- starting in 2008 because 2008 is minimum year that influenza has almost a complete set of data.

Then, removed any diseases that did not have less than or equal to 25% of monthly incidence greater than zero. This rate was choosen because it will not let an entire calendar season (fall, winter, spring, or summer) go without a single incidence. Finally, if a disease series did not have all twelve years (2008 through 2019) of data for both carrier statuses, they were removed from consideration.

# Results
```{r, echo = FALSE}
carrier_extract <- readRDS("./outputs/carrier/extract.rds")
control_extract <- readRDS("./outputs/control/extract.rds")
```

## Disease series {.tabset .tabset-fade}
```{r, include = FALSE}
col <- boonstra::pal(2)
fill <- boonstra::pal(2, alpha = 0.5)
alpha <- 0.05
crit_val <- stats::qnorm(p = 1 - (alpha / 2))
plot_names <- c(
  "Astma", "Bronchiectasis", "Acute bronchitis", "Chronic bronchitis",
  "Cough", "Influenza", "Gerd", "Nasal polyps",
  "Other lower respiratory infections", "Pneumonia", "Respiratory failure",
  "Acute sinusitis", "Chronic sinusitis",
  "Unspecified upper respiratory infections"
)
```

### Observed
```{r, echo = FALSE, out.width = "100%"}
carrier_obs <- lapply(
  X = carrier_extract,
  FUN = function(x) {
    x$obs
  }
)
control_obs <- lapply(
  X = control_extract,
  FUN = function(x) {
    x$obs
  }
)
for (j in seq_along(carrier_extract)) {
  obs <- data.frame(
    obs_carrier = carrier_obs[[j]],
    obs_control = control_obs[[j]]
  )
  graphics::par(oma = c(0, 0, 0, 7))
  astsa::tsplot(
    x = obs, main = plot_names[j], col = col,
    ylab = "Incidence per 10,000",
    xlab = "Time",
    spaghetti = TRUE
  )
  boonstra::legend_right(
    legend = c("Carriers", "Controls"), col = col, lty = 1
  )
}
```

### Trend
```{r, echo = FALSE, out.width = "100%"}
carrier_tsm <- lapply(
  X = carrier_extract,
  FUN = function(x) {
    x$tsm
  }
)
control_tsm <- lapply(
  X = control_extract,
  FUN = function(x) {
    x$tsm
  }
)
carrier_tsm_se <- lapply(
  X = carrier_extract,
  FUN = function(x) {
    x$tsm_se
  }
)
control_tsm_se <- lapply(
  X = control_extract,
  FUN = function(x) {
    x$tsm_se
  }
)
for (j in seq_along(carrier_extract)) {
  xx_carrier <- c(
    stats::time(carrier_tsm[[j]]),
    rev(stats::time(carrier_tsm[[j]]))
  )
  yy_carrier <- c(
    carrier_tsm[[j]] - (crit_val * carrier_tsm_se[[j]]),
    rev(carrier_tsm[[j]] + (crit_val * carrier_tsm_se[[j]]))
  )
  xx_control <- c(
    stats::time(control_tsm[[j]]),
    rev(stats::time(control_tsm[[j]]))
  )
  yy_control <- c(
    control_tsm[[j]] - (crit_val * control_tsm_se[[j]]),
    rev(control_tsm[[j]] + (crit_val * control_tsm_se[[j]]))
  )
  tsm <- data.frame(
    tsm_carrier = carrier_tsm[[j]],
    tsm_control = control_tsm[[j]]
  )
  ylim <- floor(
    1.10 * c(
      min(c(yy_carrier, yy_control)),
      max(c(yy_carrier, yy_control))
    )
  )
  graphics::par(oma = c(0, 0, 0, 7))
  astsa::tsplot(
    x = tsm, main = plot_names[j], col = col,
    ylab = "Incidence per 10,000", xlab = "Time",
    ylim = ylim, spaghetti = TRUE
  )
  graphics::polygon(
    x = xx_carrier, y = yy_carrier,
    col = fill[1], border = FALSE
  )
  graphics::polygon(
    x = xx_control, y = yy_control,
    col = fill[2], border = FALSE
  )
  boonstra::legend_right(
    legend = c("Carriers", "Controls"), col = col, lty = 1
  )
}
```

### Seasonal
```{r, echo = FALSE, out.width = "100%"}
carrier_ssm <- lapply(
  X = carrier_extract,
  FUN = function(x) {
    x$ssm
  }
)
control_ssm <- lapply(
  X = control_extract,
  FUN = function(x) {
    x$ssm
  }
)
carrier_ssm_se <- lapply(
  X = carrier_extract,
  FUN = function(x) {
    x$ssm_se
  }
)
control_ssm_se <- lapply(
  X = control_extract,
  FUN = function(x) {
    x$ssm_se
  }
)
for (j in seq_along(carrier_extract)) {
  xx_carrier <- c(
    stats::time(carrier_ssm[[j]]),
    rev(stats::time(carrier_ssm[[j]]))
  )
  yy_carrier <- c(
    carrier_ssm[[j]] - (crit_val * carrier_ssm_se[[j]]),
    rev(carrier_ssm[[j]] + (crit_val * carrier_ssm_se[[j]]))
  )
  xx_control <- c(
    stats::time(control_ssm[[j]]),
    rev(stats::time(control_ssm[[j]]))
  )
  yy_control <- c(
    control_ssm[[j]] - (crit_val * control_ssm_se[[j]]),
    rev(control_ssm[[j]] + (crit_val * control_ssm_se[[j]]))
  )
  ssm <- data.frame(
    ssm_carrier = carrier_ssm[[j]],
    ssm_control = control_ssm[[j]]
  )
  ylim <- floor(
    1.10 * c(
      min(c(yy_carrier, yy_control)),
      max(c(yy_carrier, yy_control))
    )
  )
  graphics::par(oma = c(0, 0, 0, 7))
  astsa::tsplot(
    x = ssm, main = plot_names[j], col = col,
    ylab = "Incidence per 10,000", xlab = "Time",
    ylim = ylim, spaghetti = TRUE
  )
  graphics::polygon(
    x = xx_carrier, y = yy_carrier,
    col = fill[1], border = FALSE
  )
  graphics::polygon(
    x = xx_control, y = yy_control,
    col = fill[2], border = FALSE
  )
  boonstra::legend_right(
    legend = c("Carriers", "Controls"), col = col, lty = 1
  )
}
```

### Local component
```{r, echo = FALSE, out.width = "100%"}
carrier_asm <- lapply(
  X = carrier_extract,
  FUN = function(x) {
    x$asm
  }
)
control_asm <- lapply(
  X = control_extract,
  FUN = function(x) {
    x$asm
  }
)
for (j in seq_along(carrier_extract)) {
  asm <- data.frame(
    obs_carrier = carrier_asm[[j]],
    obs_control = control_asm[[j]]
  )
  graphics::par(oma = c(0, 0, 0, 7))
  astsa::tsplot(
    x = asm, main = plot_names[j], col = col,
    ylab = "Incidence per 10,000",
    xlab = "Time",
    spaghetti = TRUE
  )
  boonstra::legend_right(
    legend = c("Carriers", "Controls"), col = col, lty = 1
  )
}
```